Dynamic Adaptive Saturated Sliding Mode Control For Deployment of Tethered Satellite System

### 2016-11-23 manuscript.tex
20161107
20161123

\begin{abstract}
This paper presents a novel adaptive dynamic sliding mode scheme for the deployment of the tethered satellite system. \textcolor{red}{To overcome the limited inputs of the dynamic model, a new scalable dimensionless transformation and strictly bounded terms are designed, which guarantee the existence of the ratio between limited inputs and command signals.} The adaptive saturated sliding mode control scheme is proposed to govern the tether deployment by introducing the adaptive rate into the system gain. The stability analysis indicates that this measure can eliminate the uncertainty caused by the input limitation and guarantee the asymptotically stability of the deployment dynamics. \textcolor{red}{Finally, numerical simulations are presented to verify the effectiveness of the proposed control laws.}
\end{abstract}
\begin{keyword}
Tethered satellite system;  Sliding mode control;  Input limitation; Adaptive control
\end{keyword}
\end{frontmatter}
\section{Introduction}
The deployment is the key technology of the space tethered satellite system (TSS)~\cite{wen2008advances,yousefian2015anti}, which supports various applications for space explorations and scientific researches, such as space debris removals, space elevators, satellite de-orbits and the tethered space robots~\cite{zhao2014thrust,kojima2015mission,zhong2013dynamics,zhong2013long,huang2014optimal,Huang2016Impact,Huang201637,Huang2016}. \textcolor{red}{Many international space research institutes have launched orbital, suborbital and ground experiments, such as TSS-1R, SEDS, MAST and YES2, in an effort to verify the feasibility of the proposed theoretical and simulation schemes for the deployment of the TSS~\cite{williams2012review,robitaille2006interpreting,stone1998tss,williams2009yes2}. These experimental results intensively help researchers proceed to promote advanced mechanics, kinetic and control studies, to tackle the future challenges of space activities~\cite{yu2016nonlinear,wen2015space,aslanov2016swing,meng2016lyapunov,zhang2015line,Xu2016Hybrid}.}\par

\textcolor{red}{Recent scientific researches about deployments of the TSS focus on the stabilization of the dynamics, and have provided different kinds of control schemes, including tension, thruster, electrodynamic and Coulomb approaches~\cite{kojima2015stabilisation,qi2016dynamics,zhong2016research}.} Among these stabilization strategies, the hybrid tension and thruster control technique has been a great boost to new developments of the deployment of the TSS. Iki et al. implemented the ground deployment experiment, and utilized the thruster to conduct the deployment of a kilometers-length electrodynamic tether from a special spool-type reel~\cite{iki2014experiments}. \textcolor{red}{Aslanov et al. presented a novel model about the spatial motion of a large passive satellite and the tug connected by space tether, and the motion of the overall system was excited by a thruster force acting on the tug~\cite{aslanov2013dynamics}.} Huang~et~al. investigated the attitude and deployment dynamics of the tethered space robot from the standpoint of thruster consumption~\cite{huang2015coupling}. Jung et al. adopted the thruster scheme to achieve the ideal deployment rate during the deployment of three-body tethered satellites~\cite{jung2015nonlinear}. \textcolor{red}{Liu et al. equipped the subsatellite with the thruster system, consisting of four pairs of thrusters, in order to regulate the attitude and deployment dynamics~\cite{liu2014attitude}. Zakrzhevskii divided the deployment process into two stages: in the first stage, the thruster was used to propel the subsatellite to the desired starting point of the next stage, and in the second stage, the tether was adjusted to the stable vertical configuration using tension generated by friction~\cite{zakrzhevskii2016method}. Mantellato et al. reported the thruster-aided deployment of tape tethers driven by both the tension and thruster forces~\cite{mantellato2015thrust}. These hybrid strategies has several advantages, such as simplifying the input model greatly, non-environmental dependence (magnetic field dependence), easy assembly and physical implementation, and thus is broadly utilized in engineering applications compared with other control strategies.}\par

However, there also exist some obstacles in the hybrid control design. Particularly, it's a prime problem that tension actuators and thrusters are both constrained by physical and chemical conditions, i.e., the space tether is not available to be repressed like the stick, and meanwhile too big tether length rate might break the tether. Also, the thruster can not provide the infinite propelling force~\cite{martinez1998spacecraft}. In simpler terms, the tension and the thruster force for the deployment are both limited, and in the view of control theory, the phenomena usually causes the input limitation of a dynamic system. Substituting limited inputs into unmodified conventional control scheme designed for systems with unlimited inputs results in poor performance or instability, and theoretically, such kind of implement is unreasonable. Hence to cope with the issues mentioned above, many scientific researchers spent much effort on the issue about the input limitation, providing many solutions on satellite arrangements and deployments. Huang et al. proposed an adaptive postcapture backstepping control law for the tethered space robot, and analyzed the effect from constrained inputs by using specific auxiliary design systems~\cite{huang2015adaptive}. Wen et al. took into account the positive tension constraint and designed the feedback control with the special saturation technique~\cite{wen2016constrained}. Wen also proposed some remarkable achievements about deployment dynamics using the optimal control theory~\cite{wen2008optimal}. Liu et al. reported the variable structure scheme for the deployment with the special structural thruster system, where the thrusters were used to ensure the positive tension during the mission~\cite{yingying2012variable}. In summary, existing studies deal with the major issue about the deployment with restricted tension and thruster force in two ways, accordingly, the suitable physical compensation and the elaborate design of limited input. In aspect of designing elaborate limited inputs, as Wen et al. explained, the limited terms, including the tension and the states of deployment dynamics, can be viewed as seeking the solution of a nonlinear optimal control issue~\cite{wen2015tension}. \textcolor{red}{Hu reported an adaptive sliding mode control law with the input limitation by using the reduced sliding dynamic design approach~\cite{Hu2009Robust}, which requires an assumption that the command signal is bounded strictly, and for actual physical systems, this assumption always exits.} In this paper, a new dynamic adaptive saturated sliding mode  control for the deployment of the TSS is studied. For coping with the input limitation in the hybrid control strategy, the strictly saturated command signal is constructed with Sine saturation function. Based on the feature of the saturated command signal, an adaptive rate is introduced into the control law design to remove the negative effect of the input limitation theoretically. In order to improve the performance of the sliding phase, the terminal attractor is incorporated into the design of the ideal sliding surface. We perform numerical simulations to illustrate that the stabilization of the deployment is accomplished by the control law designed in spite of the input limitation.\par

@@

\textcolor{red}{Figure \ref{fig:1} shows all the elements of the TSS via the geometric representation. In the figure, the axes of the coordinate frame $OXYZ$ passes through the center of the earth $O$. $OX$ and $OY$ are mutually perpendicular axes on the equatorial plane with $OX$ pointing to the spring equinox, and $OZ$ is perpendicular to the equatorial plane, bundling the axis of the earth upwards. The axis $OY$ completes the right-hand triad. The orbital coordinate frame is denoted as $O_2xyz$ with the origin $O_2$ located at the mass center of the TSS. $O_2x$ is directed in the direction of orbital velocity, meanwhile $O_2z$ is defined to point towards the center of the earth, generating the axis $O_2y$ by utilizing the right-hand triad. The origin of the TSS coordinate frame $O_2x'y'z'$ is also fixed at the mass center of the TSS, and the axis $O_2z'$ is directed along the tether and points to the subsatellite. The orientation of $O_2xyz$ relative to the coordinate frame $O_2x'y'z'$ indicates the in-plane and out-of-plane dynamics, where $\theta$ and $\phi$ stand for the in-plane angle and out-of-plane angle, respectively. $\bm R_0$ denotes the position vector between $O_2$ and $O$.}\par
\textcolor{red}{In the analysis of the deployment of the TSS, there is a primary assumption that the movement of entire system is similar to a ``rolling ball'' behavior with three DOFs modelled, and it is reasonable to invoke the spherical coordinates to describe the behavior~\cite{Williams2008}.} During space mission, the mass of the tether is much smaller than that of the mother satellite, and hence it is always ignored sufficiently. Some other assumptions are obvious and as follows. \textcolor{red}{The sizes of satellites are small compared with the length of the tether, and hence a satellite can be regard as a particle if the tether is long enough. Actually, the mass ratio between the subsatellite and the mother satellite is tiny, i.e., in YES2, the mother satellite mass is 6530~$kg$, and the subsatellite mass is 12~$kg$~\cite{Williams2008}.  Owing to the tiny mass ratio between the satellites, the deployment of the subsatellite will not affect the motion of the mother satellite during the whole mission.} The deployment is regulated by the tension and thrusters, and to guarantee the establishment of the ``rolling ball'' model, the tether remains inextensible, straight and stiff all the time.\par

@@

&\bar m l''-\bar m l\begin{bmatrix}\phi'^2+(\theta'+\Omega)^2\cos^2\phi\end{bmatrix}+\bar m l\Omega^2(1-3\cos^2\phi\cos^2\theta)=-\tau_t+\tau_{td}\\

@@

&\bar m l^2\phi''+2\bar m \phi'l'l+\bar m l^2\sin\phi\cos\phi\begin{bmatrix}(\theta'+\Omega)^2+3\bar m \Omega\cos^2\theta\end{bmatrix}=\tau_\phi+\tau_{\phi d},\label{eq:dynamics 1}

@@

\textcolor{red}{where, $m = m_1+m_2$ stands for the total mass of the system, $m_1$ and $m_2$ of which denote the mass of the mother satellite and of the subsatellite, respectively. \textcolor{red}{$(\cdot)'$ represents a first derivative with respect to time, and $l$ means the tether length connecting the satellites. $\theta$ and $\phi$ depict the roll and pitch dynamics of the TSS. For clarity, $m_1m_2/m$ is rewritten as $\bar m$. $\tau_t$, $\tau_\theta$ and $\tau_\phi$ are the tension, the in-plane thruster force and the out-of-plane thruster force, respectively. $\tau_{t d}$, $\tau_{\theta d}$ and $\tau_{\phi d}$ stand for the bounded external disturbances correspondingly.} Typically, the dimensionless transformations simplify issues for researchers to design the deployment scheme, and the conventional transformations are represented as shown below~\cite{williams2009yes2,wen2015space,wen2016constrained}:

@@

&d()/dt=\Omega d()/dv\\

@@

The obvious benefits of these transformations are that all the inputs own the positive and negative boundedness and hence satisfy the assumption about limited inputs of existing control laws~\cite{Hu2008552,6060930}. Furthermore, the generalized inputs~(\ref{eq:varcon}) can be processed by the proposed Sine function in this paper to yield the strict saturated inputs, which essentially guarantee the rigorous stability analysis of the proposed adaptive sliding mode control law.}

@@

\textcolor{red}{In reality, once $\lambda = 0$ is satisfied, the mother satellite and subsatellite will collide with each other, which is meaningless for the deployment of the TSS. Hence the minimum of the dimensionless length $\lambda_{min}$ is introduced to maintain the reasonably safe formation of the satellites, and also guarantees the non-singularity of the dimensionless transformation~(\ref{eq:dynamics 1}) theoretically.} $\lambda_{max}$ stands for the maximum tether length ratio.\par

@@

\textcolor{red}{For the system~(\ref{eq:dynamic model 2}) with the input limitation ignored, the author in \cite{parra2003dynamic} has designed a nonlinear dynamic sliding PID controller with the zero initial condition:

@@

\dot{\bm\sigma}&=\bm{sign}(\bm S)\\

@@

\bm{sign}(\bm x) = \left(sign(x_1),sign(x_2),\cdots,sign(x_n)\right)^T

@@

for a vector $\bm x=\left(x_1,x_2,,\cdots,x_n\right)^T$. $\alpha$ and $\gamma$ are control law parameters which will be discussed in the control law design. Define the tracking error as $\Delta \bm q=\bm q-\bm q_d$, in which $\bm q_d$ is the desired trajectory with $\bm q_d(t)\in C^2$. However, the PID controller~(\ref{eq:pid controller}) can not be directly applied to the system~(\ref{eq:dynamic model 2}) with the input limitation, and in the subsequent section, we will utilize the saturated sliding mode technique to modify the controller~(\ref{eq:pid controller}) for the system~(\ref{eq:dynamic model 2}).} For deriving the proposed control scheme, some properties should be given out in advance.

@@

\textcolor{red}{Positive scalars $\Xi_i(i = 0,\ldots,6)$ always exist such that

@@

where $\zeta_{min}(\cdot)$ and $\zeta_{max}(\cdot)$ mean the minimum and maximum eigenvalues of matrix, respectively. Furthermore, one can have $Y_r\bm\Theta\le \eta(t)$, in which $\eta(t)=f(\Delta \bm q,\Delta \dot {\bm q},\sigma,\Xi_i,t)$ is a state-dependent function, thus $Y_r\bm\Theta$ is usually thought to be a bounded function with $\vert\eta\vert\le\varphi_0$, where $\varphi_0$ is a positive scalar determined according to the system states.}

@@

\textcolor{red}{

@@

}

@@

Sine(x) =

@@

sign(x)&\vert x\vert\ge\pi/2\\
sin(x)&\vert x\vert<\pi/2

@@

and it's notable that the function above is saturated and owns an important property: $\vert Sine(x)\vert\le\vert x\vert$, $\forall x$. The geometric illustration is shown in Figure~\ref{fig:2}, from which an interesting property can be found by sliding the curve $y=Sine(x)$ along the axis $x=0$ in the range $[-1,1]$, and once the function $y=Sine(x)+\Delta h$ with $\vert\Delta h\vert<1$ settles, one can choose the function $y = x+arcsin(\Delta h)$ to obtain following relations:

@@

&\vert Sine(x)+\Delta h\vert\le\vert x+arcsin(\Delta h)\vert\label{eq:lem11}\\
&sign(Sine(x)+\Delta h)=sign(x+arcsin(\Delta h)), \forall x, \vert\Delta h\vert<1\label{eq:lem12}.

@@

For relation~(\ref{eq:lem11}), let $y=Sine(x)+\Delta h-x-arcsin(\Delta h)$, and we can get its derivative

@@

dy/dx = -1,&x\ge\pi/2,x\le-\pi/2\\
dy/dx = \cos x - 1,&\vert x\vert<\pi/2,

@@

where $dy/dx<0$ means $y$ is a monotone decreasing function with respect to $x$. As $x=-arcsin(\Delta h)$, $y=0$, according to the definition of $Sine(x)$, we proceed to get

@@

x\ge-arcsin(\Delta h), y\le0&\Rightarrow 0\le Sine(x)+\Delta h\le x+arcsin(\Delta h)\\
x<-arcsin(\Delta h), y>0&\Rightarrow 0> Sine(x)+\Delta h>x+arcsin(\Delta h)\end{aligned}\right\}\\\\
\Rightarrow\vert Sine(x)+\Delta h\vert\le\vert x+arcsin(\Delta h)\vert.

@@

For Eq.~(\ref{eq:lem12}), considering the case $Sine(x)+\Delta h>0$, when $\vert x\vert<\pi/2$, $Sine(x)+\Delta h>0$ becomes

@@

sin(x)+\Delta h>0&\Rightarrow x>-arcsin(\Delta h)\\
&\Rightarrow x+arcsin(\Delta h)>0,

@@

since $y=arcsin(x)$ is a monotone increasing function. We only need to consider $x\ge\pi/2$ instead of $\vert x\vert\ge\pi/2$ for $x\le-\pi/2$ is meaningless in the case $Sine(x)+\Delta h>0$. It is obvious that $x+arcsin(\Delta h)>0$ due to $\vert arcsin(\Delta h)\vert<\pi/2$. Therefore, we can get

@@

sign(Sine(x)+\Delta h)=sign(x+arcsin(\Delta h)),Sine(x)+\Delta h>0,

@@

sign(Sine(x)+\Delta h)=sign(x+arcsin(\Delta h)),Sine(x)+\Delta h\le0.

@@

\begin{myrem}\textcolor{red}{Moreover, we think that a quantity $y$ is saturated iff it is produced directly by the saturated function $y=Sine(x)$. Otherwise, the bounded quantity is different from the saturated one since it does not own the standard formulation like $y=Sine(x)$, i.e., a quantity can just be referred to as a bounded quantity if its boundary is $[a,b]$ with $a\neq -1$ or $b\neq 1$.}

@@

\caption{$y =Sine(x)$ and $y=Sine(x)+\Delta h$}

@@

As we know, \textcolor{red}{the tension, in-plane and out-of-plane thruster forces are all limited during the deployment mission, and we describe these inputs by using the matrix and vector to predigest the following illation.}

@@

and in order to design the saturated sliding mode control law for the dynamics of the deployment, the $Sine$ function for vectors is specified as:

@@

\bm{Sine}(\bm x)=\begin{pmatrix}Sine(x_1),Sine(x_2),\cdots,Sine(x_n)\end{pmatrix}^T.

@@

\bm\tau = -\bm{g}(\delta\varphi F(t) \bm{sign}(\bm S_{r1})),\label{eq:slider control}

@@

\bm S_{q1} &= \bm{Sine}(\Delta \dot {\bm q}+\alpha\Delta\bm q),

@@

\frac{d}{dt}F(t) = \delta\varphi \bm S_{r2}^T K_1e^{-2K_1^{-1}F^{-1}(t)}\bm{sign}(\bm S_{r2})

@@

\textcolor{red}{where $K_1\in R^{3\times 3}$ is a positive definite diagonal matrix, which will be given out later, and if $K\in R^{3\times 3}$, define}

@@

e^K=

@@

e^{K_{11}}&e^{K_{12}}&e^{K_{13}}\\
e^{K_{21}}&e^{K_{22}}&e^{K_{23}}\\
e^{K_{31}}&e^{K_{32}}&e^{K_{33}}

@@

\textcolor{red}{$\varphi$ is the boundedness of the lumped uncertainty. $\gamma$ is designed as a time-varying positive scalar and $\delta>1$ is the design parameter, so that the deployment dynamics can be regulated as desired.}

@@

\bm S_{r2} &= \bm S_{q2}+\bm{arcsin}(\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) d\varsigma)\\

@@

where $\bm y=\bm{arcsin}(\bm x)$ is the vector function with

>>>>

\begin{abstract}
This paper presents a novel adaptive dynamic sliding mode scheme for the deployment of the tethered satellite system. To overcome the limited inputs of the dynamic model, a new scalable dimensionless transformation and strictly bounded terms are designed, which guarantee the existence of the ratio between limited inputs and command signals. The adaptive saturated sliding mode control scheme is proposed to govern the tether deployment by introducing the adaptive rate into the system gain. The stability analysis indicates that this measure can eliminate the uncertainty caused by the input limitation and guarantee the asymptotically stability of the deployment dynamics. Finally, numerical simulations are presented to verify the effectiveness of the proposed control laws.
\end{abstract}
\begin{keyword}
Tethered satellite system;  Sliding mode control;  Input limitation; Adaptive control
\end{keyword}
\end{frontmatter}
\section{Introduction}
The deployment is the key technology of the space tethered satellite system (TSS)~\cite{wen2008advances,yousefian2015anti}, which supports various applications for space explorations and scientific researches, such as space debris removals, space elevators, satellite de-orbits and the tethered space robots~\cite{zhao2014thrust,kojima2015mission,zhong2013dynamics,zhong2013long,huang2014optimal,Huang2016Impact,Huang201637,Huang2016}. Many international space research institutes have launched orbital, suborbital and ground experiments, such as TSS-1R, SEDS, MAST and YES2, in an effort to verify the feasibility of the proposed theoretical and simulation schemes for the deployment of the TSS~\cite{williams2012review,robitaille2006interpreting,stone1998tss,williams2009yes2}. These experimental results intensively help researchers proceed to promote advanced mechanics, kinetic and control studies, to tackle the future challenges of space activities~\cite{yu2016nonlinear,wen2015space,aslanov2016swing,meng2016lyapunov,zhang2015line,Xu2016Hybrid}.

Recent scientific researches about deployments of the TSS focus on the stabilization of the dynamics, and have provided different kinds of control schemes, including tension, thruster, electrodynamic and Coulomb approaches~\cite{kojima2015stabilisation,qi2016dynamics,zhong2016research}. Among these stabilization strategies, the hybrid tension and thruster control technique has been a great boost to new developments of the deployment of the TSS. Iki et al. implemented the ground deployment experiment, and utilized the thruster to conduct the deployment of a kilometers-length electrodynamic tether from a special spool-type reel~\cite{iki2014experiments}. Aslanov et al. presented a novel model about the spatial motion of a large passive satellite and the tug connected by space tether, and the motion of the overall system was excited by a thruster force acting on the tug~\cite{aslanov2013dynamics}. Huang~et~al. investigated the attitude and deployment dynamics of the tethered space robot from the standpoint of thruster consumption~\cite{huang2015coupling}. Jung et al. adopted the thruster scheme to achieve the ideal deployment rate during the deployment of three-body tethered satellites~\cite{jung2015nonlinear}. Liu et al. equipped the subsatellite with the thruster system, consisting of four pairs of thrusters, in order to regulate the attitude and deployment dynamics~\cite{liu2014attitude}. Zakrzhevskii divided the deployment process into two stages: in the first stage, the thruster was used to propel the subsatellite to the desired starting point of the next stage, and in the second stage, the tether was adjusted to the stable vertical configuration using tension generated by friction~\cite{zakrzhevskii2016method}. Mantellato et al. reported the thruster-aided deployment of tape tethers driven by both the tension and thruster forces~\cite{mantellato2015thrust}. These hybrid strategies has several advantages, such as simplifying the input model greatly, non-environmental dependence (magnetic field dependence), easy assembly and physical implementation, and thus is broadly utilized in engineering applications compared with other control strategies.

However, there also exist some obstacles in the hybrid control design. Particularly, it's a prime problem that tension actuators and thrusters are both constrained by physical and chemical conditions, i.e., the space tether is not available to be repressed like the stick, and meanwhile too big tether length rate might break the tether. Also, the thruster can not provide the infinite propelling force~\cite{martinez1998spacecraft}. In simpler terms, the tension and the thruster force for the deployment are both limited, and in the view of control theory, the phenomena usually causes the input limitation of a dynamic system. Substituting limited inputs into unmodified conventional control scheme designed for systems with unlimited inputs results in poor performance or instability, and theoretically, such kind of implement is unreasonable. Hence to cope with the issues mentioned above, many scientific researchers spent much effort on the issue about the input limitation, providing many solutions on satellite arrangements and deployments. Huang et al. proposed an adaptive postcapture backstepping control law for the tethered space robot, and analyzed the effect from constrained inputs by using specific auxiliary design systems~\cite{huang2015adaptive}. Wen et al. took into account the positive tension constraint and designed the feedback control with the special saturation technique~\cite{wen2016constrained}. Wen also proposed some remarkable achievements about deployment dynamics using the optimal control theory~\cite{wen2008optimal}. Liu et al. reported the variable structure scheme for the deployment with the special structural thruster system, where the thrusters were used to ensure the positive tension during the mission~\cite{yingying2012variable}. In summary, existing studies deal with the major issue about the deployment with restricted tension and thruster force in two ways, accordingly, the suitable physical compensation and the elaborate design of limited input. In aspect of designing elaborate limited inputs, as Wen et al. explained, the limited terms, including the tension and the states of deployment dynamics, can be viewed as seeking the solution of a nonlinear optimal control issue~\cite{wen2015tension}. Hu reported an adaptive sliding mode control law with the input limitation by using the reduced sliding dynamic design approach~\cite{Hu2009Robust}, which requires an assumption that the command signal is bounded strictly, and for actual physical systems, this assumption always exits. In this paper, a new dynamic adaptive saturated sliding mode  control for the deployment of the TSS is studied. For coping with the input limitation in the hybrid control strategy, the strictly saturated command signal is constructed with Sine saturation function. Based on the feature of the saturated command signal, an adaptive rate is introduced into the control law design to remove the negative effect of the input limitation theoretically. In order to improve the performance of the sliding phase, the terminal attractor is incorporated into the design of the ideal sliding surface. We perform numerical simulations to illustrate that the stabilization of the deployment is accomplished by the control law designed in spite of the input limitation.\par

@@

Figure \ref{fig:1} shows all the elements of the TSS via the geometric representation. In the figure, the axes of the coordinate frame $OXYZ$ passes through the center of the earth $O$. $OX$ and $OY$ are mutually perpendicular axes on the equatorial plane with $OX$ pointing to the spring equinox, and $OZ$ is perpendicular to the equatorial plane, bundling the axis of the earth upwards. The axis $OY$ completes the right-hand triad. The orbital coordinate frame is denoted as $O_2xyz$ with the origin $O_2$ located at the mass center of the TSS. $O_2x$ is directed in the direction of orbital velocity, meanwhile $O_2z$ is defined to point towards the center of the earth, generating the axis $O_2y$ by utilizing the right-hand triad. The origin of the TSS coordinate frame $O_2x'y'z'$ is also fixed at the mass center of the TSS, and the axis $O_2z'$ is directed along the tether and points to the subsatellite. The orientation of $O_2xyz$ relative to the coordinate frame $O_2x'y'z'$ indicates the in-plane and out-of-plane dynamics, where $\theta$ and $\phi$ stand for the in-plane angle and out-of-plane angle, respectively. $\bm R_0$ denotes the position vector between $O_2$ and $O$.

In the analysis of the deployment of the TSS, there is a primary assumption that the movement of entire system is similar to a ``rolling ball'' behavior with three DOFs modelled, and it is reasonable to invoke the spherical coordinates to describe the behavior~\cite{Williams2008}. During space mission, the mass of the tether is much smaller than that of the mother satellite, and hence it is always ignored sufficiently. Some other assumptions are obvious and as follows. The sizes of satellites are small compared with the length of the tether, and hence a satellite can be regard as a particle if the tether is long enough. Actually, the mass ratio between the subsatellite and the mother satellite is tiny, i.e., in YES2, the mother satellite mass is 6530~$kg$, and the subsatellite mass is 12~$kg$~\cite{Williams2008}.  Owing to the tiny mass ratio between the satellites, the deployment of the subsatellite will not affect the motion of the mother satellite during the whole mission. The deployment is regulated by the tension and thrusters, and to guarantee the establishment of the ``rolling ball'' model, the tether remains inextensible, straight and stiff all the time.\par

@@

&\bar m l''-\bar m l\left\{\begin{bmatrix}\phi'^2+(\theta'+\Omega)^2\cos^2\phi\end{bmatrix}+\Omega^2(3\cos^2\phi\cos^2\theta-1)\right\}=-\tau_t+\tau_{td}\\

@@

&\bar m l^2\phi''+2\bar m \phi'l'l+\bar m l^2\sin\phi\cos\phi\left[(\theta'+\Omega)^2+3\bar m \Omega\cos^2\theta\right]=\tau_\phi+\tau_{\phi d},\label{eq:dynamics 1}

@@

where, $m = m_1+m_2$ stands for the total mass of the system, $m_1$ and $m_2$ of which denote the mass of the mother satellite and of the subsatellite, respectively. $(\cdot)'$ represents a first derivative with respect to time, and $l$ means the tether length connecting the satellites. $\theta$ and $\phi$ depict the roll and pitch dynamics of the TSS. For clarity, $m_1m_2/m$ is rewritten as $\bar m$. $\tau_t$, $\tau_\theta$ and $\tau_\phi$ are the tension, the in-plane thruster force and the out-of-plane thruster force, respectively. $\tau_{t d}$, $\tau_{\theta d}$ and $\tau_{\phi d}$ stand for the bounded external disturbances correspondingly. Typically, the dimensionless transformations simplify issues for researchers to design the deployment scheme, and the conventional transformations are represented as shown below~\cite{williams2009yes2,wen2015space,wen2016constrained}:

@@

&\rm{d}\it{()}/\rm{d}\it{t}=\Omega \rm{d}\it{()}/\rm{d}\it{v}\\

@@

The obvious benefits of these transformations are that all the inputs own the positive and negative boundedness and hence satisfy the assumption about limited inputs of existing control laws~\cite{Hu2008552,6060930}. Furthermore, the generalized inputs~(\ref{eq:varcon}) can be processed by the proposed Sine function in this paper to yield the strict saturated inputs, which essentially guarantee the rigorous stability analysis of the proposed adaptive sliding mode control law.

@@

In reality, once $\lambda = 0$ is satisfied, the mother satellite and subsatellite will collide with each other, which is meaningless for the deployment of the TSS. Hence the minimum of the dimensionless length $\lambda_{min}$ is introduced to maintain the reasonably safe formation of the satellites, and also guarantees the non-singularity of the dimensionless transformation~(\ref{eq:dynamics 1}) theoretically. $\lambda_{max}$ stands for the maximum tether length ratio.\par

@@

For the system~(\ref{eq:dynamic model 2}) with the input limitation ignored, the author in \cite{parra2003dynamic} has designed a nonlinear dynamic sliding PID controller with the zero initial condition:

@@

\dot{\bm\sigma}&=\mathbf{sign}(\bm S)\\

@@

\mathbf{sign}(\bm x) = \left(\rm{sign}(\it{x_1}),\rm{sign}(\it{x_2}),\cdots,\rm{sign}(\it{x_n})\right)^T

@@

for a vector $\bm x=\left(x_1,x_2,,\cdots,x_n\right)^T$. $\alpha$ and $\gamma$ are control law parameters which will be discussed in the control law design. Define the tracking error as $\Delta \bm q=\bm q-\bm q_d$, in which $\bm q_d$ is the desired trajectory with $\bm q_d(t)\in C^2$. However, the PID controller~(\ref{eq:pid controller}) can not be directly applied to the system~(\ref{eq:dynamic model 2}) with the input limitation, and in the subsequent section, we will utilize the saturated sliding mode technique to modify the controller~(\ref{eq:pid controller}) for the system~(\ref{eq:dynamic model 2}). For deriving the proposed control scheme, some properties should be given out in advance.

@@

Positive scalars $\Xi_i(i = 0,\ldots,6)$ always exist such that

@@

where $\zeta_{min}(\cdot)$ and $\zeta_{max}(\cdot)$ mean the minimum and maximum eigenvalues of matrix, respectively. Furthermore, one can have $Y_r\bm\Theta\le \eta(t)$, in which $\eta(t)=f(\Delta \bm q,\Delta \dot {\bm q},\sigma,\Xi_i,t)$ is a state-dependent function, thus $Y_r\bm\Theta$ is usually thought to be a bounded function with $\vert\eta\vert\le\varphi_0$, where $\varphi_0$ is a positive scalar determined according to the system states.

@@

@@

@@

\rm{Sine}\it{(x)} =

@@

\rm{sign}\it{(x)}&\vert x\vert\ge\pi/2\\
\sin(x)&\vert x\vert<\pi/2

@@

and it's notable that the function above is saturated and owns an important property: $\vert \rm{Sine}\it{(x)}\vert\le\vert x\vert$, $\forall x$. The geometric illustration is shown in Figure~\ref{fig:2}, from which an interesting property can be found by sliding the curve $y=\rm{Sine}(\it{x})$ along the axis $x=0$ in the range $[-1,1]$, and once the function $y=\rm{Sine}\it{(x)}+\Delta h$ with $\vert\Delta h\vert<1$ settles, one can choose the function $y = x+\arcsin(\Delta h)$ to obtain following relations:

@@

&\vert \rm{Sine}\it{(x)}+\Delta h\vert\le\vert x+\arcsin(\Delta h)\vert\label{eq:lem11}\\
&\rm{sign}(\rm{Sine}\it{(x)}+\Delta h)=\rm{sign}(\it{x}+\arcsin(\Delta h)), \forall x, \vert\Delta h\vert<1\label{eq:lem12}.

@@

For relation~(\ref{eq:lem11}), let $y=\rm{Sine}\it{(x)}+\Delta h-x-\arcsin(\Delta h)$, and we can get its derivative

@@

\rm{d}\it{y}/\rm{d}\it{x} = -1,&x\ge\pi/2,x\le-\pi/2\\
\rm{d}\it{y}/\rm{d}\it{x} = \cos x - 1,&\vert x\vert<\pi/2,

@@

where $\rm{d}\it{y}/\rm{d}\it{x}<0$ means $y$ is a monotone decreasing function with respect to $x$. As $x=-\arcsin(\Delta h)$, $y=0$, according to the definition of $\rm{Sine}\it{(x)}$, we proceed to get

@@

x\ge-\arcsin(\Delta h), y\le0&\Rightarrow 0\le \rm{Sine}\it{(x)}+\Delta h\le x+\arcsin(\Delta h)\\
x<-\arcsin(\Delta h), y>0&\Rightarrow 0> \rm{Sine}\it{(x)}+\Delta h>x+\arcsin(\Delta h)\end{aligned}\right\}\\\\
\Rightarrow\vert \rm{Sine}\it{(x)}+\Delta h\vert\le\vert x+\arcsin(\Delta h)\vert.

@@

For Eq.~(\ref{eq:lem12}), considering the case $\rm{Sine}\it{(x)}+\Delta h>0$, when $\vert x\vert<\pi/2$, $\rm{Sine}\it{(x)}+\Delta h>0$ becomes

@@

\sin(x)+\Delta h>0&\Rightarrow x>-\arcsin(\Delta h)\\
&\Rightarrow x+\arcsin(\Delta h)>0,

@@

since $y=\arcsin(x)$ is a monotone increasing function. We only need to consider $x\ge\pi/2$ instead of $\vert x\vert\ge\pi/2$ for $x\le-\pi/2$ is meaningless in the case $\rm{Sine}\it{(x)}+\Delta h>0$. It is obvious that $x+\arcsin(\Delta h)>0$ due to $\vert \arcsin(\Delta h)\vert<\pi/2$. Therefore, we can get

@@

\rm{sign}(\rm{Sine}\it{(x)}+\Delta h)=\rm{sign}\it{(x}+\arcsin(\Delta h)),\rm{Sine}\it{(x)}+\Delta h>0,

@@

\rm{sign}(\rm{Sine}\it{(x)}+\Delta h)=\rm{sign}\it{(x}+\arcsin(\Delta h)),\rm{Sine}\it{(x)}+\Delta h\le0.

@@

\begin{myrem}Moreover, we think that a quantity $y$ is saturated iff it is produced directly by the saturated function $y=\rm{Sine}\it{(x)}$. Otherwise, the bounded quantity is different from the saturated one since it does not own the standard formulation like $y=\rm{Sine}\it{(x)}$, i.e., a quantity can just be referred to as a bounded quantity if its boundary is $[a,b]$ with $a\neq -1$ or $b\neq 1$.

@@

\caption{$y =\rm{Sine}\it{(x)}$ and $y=\rm{Sine}\it{(x)}+\Delta h$}

@@

As we know, the tension, in-plane and out-of-plane thruster forces are all limited during the deployment mission, and we describe these inputs by using the matrix and vector to predigest the following illation.

@@

and in order to design the saturated sliding mode control law for the dynamics of the deployment, the $\rm{Sine}$ function for vectors is specified as:

@@

\mathbf{Sine}(\bm x)=\begin{pmatrix}\rm{Sine}\it{(x_1)},\rm{Sine}\it{(x_2)},\cdots,\rm{Sine}\it{(x_n)}\end{pmatrix}^T.

@@

\bm\tau = -\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r1})),\label{eq:slider control}

@@

\bm S_{q1} &= \mathbf{Sine}(\Delta \dot {\bm q}+\alpha\Delta\bm q),

@@

\frac{\rm{d}}{\rm{d}\it{t}}F(t) = \delta\varphi \bm S_{r2}^T K_1\rm{e}\it{^{-2K_1^{-1}F^{-1}(t)}}\mathbf{sign}(\bm S_{r2})

@@

where $K_1\in R^{3\times 3}$ is a positive definite diagonal matrix, which will be given out later, and if $K\in R^{3\times 3}$, define

@@

\rm{e}\it{^K}=

@@

\rm{e}\it{^{K_{11}}}&\rm{e}\it{^{K_{12}}}&\rm{e}\it{^{K_{13}}}\\
\rm{e}\it{^{K_{21}}}&\rm{e}\it{^{K_{22}}}&\rm{e}\it{^{K_{23}}}\\
\rm{e}\it{^{K_{31}}}&\rm{e}\it{^{K_{32}}}&\rm{e}\it{^{K_{33}}}

@@

$\varphi$ is the boundedness of the lumped uncertainty. $\gamma$ is designed as a time-varying positive scalar and $\delta>1$ is the design parameter, so that the deployment dynamics can be regulated as desired.

@@

\bm S_{r2} &= \bm S_{q2}+\bm{arcsin}(\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma})\\

@@

where $\bm y=\mathbf{arcsin}(\bm x)$ is the vector function with
